---
title: "MCP for TCGA computations with Bioconductor"
shorttitle: "outlier dytection methods"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{MCP for TCGA computations with Bioconductor}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

We want to enable a reliable natural language interface to
curated TCGA facilities provided by Bioconductor.  One way to
do this is through creation and registration of MCP-compliant tools
that use Bioconductor facilities to prepare conversational
responses and data objects to active R sessions.

# Example

In this example we require that a value has been provided for
environment variable `OPENAI_API_KEY`.  We also assume that
MCP configuration has been accomplished for the host system.
Typically this involves a schema placed in `~/.config/mcptools`

## Unstructured responses

### Cancer type enumeration with and without MCP
Enumerate cancer types:

```{r doone, message=FALSE}
library(ellmer)
library(biocTcgaMcp)
if (nchar(Sys.getenv("OPENAI_API_KEY"))>0) {
ch = chat_openai(model="gpt-4o")
ch$chat("List the tumor types in TCGA")
}
```

The above result can have more or less than 31 cancer types.  To
assert control over the operation

```{r setuptools}
if (nchar(Sys.getenv("OPENAI_API_KEY"))>0) {
ch$set_tools(mcptools::mcp_tools())
ch$chat("List the tumor types in TCGA")
}
```

### Assay enumeration for a tumor type.


```{r domore}
if (nchar(Sys.getenv("OPENAI_API_KEY"))>0) {
ch$chat("What assays are available for lung adenocarcinoma? Please provide the
answer in a markdown table.")
}
```

## Structured response

We can use the `chat_structured` facility to return data directly
to the R session.  `cantypes_str` is a hand-coded schema prepared
for cancer type enumeration in a list.  Here we reset the model
in use.

```{r dostruc}
if (nchar(Sys.getenv("OPENAI_API_KEY"))>0) {
ch = chat_openai(model="gpt-4o")
ch$set_tools(mcptools::mcp_tools())
ans = ch$chat_structured("List the tumor types in TCGA that available assays in curatedTCGAData.", type=cantypes_str())
DT::datatable(ans)
}
```
